Generating exact lattices in a WFST framework

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Removed some unwanted stuff from lattice paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@511 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

We report experiments on CallHome English and Switchboard.

Our Callhome English setup is as described in~\cite{ubmconf,ubmpaper}.  
We used PLP features with ceptral mean and variance normalization.  We tested 
with the trigram LM built as described in~\cite{ubmpaper}.

%% [dan's results, not included, used arnab's.]
%%svatava:er: showwer_nist test_arnab_nosmooth_floor02_vecs_rerun/*/*/*0.0.ctm.filt.dtl
%%3/4:48.7% 3/8:48.3% 4/4:47.9% 4/8:47.5% 5/4:47.3% 5/8:47.2% 6/4:47.2% 6/8:46.9% 7/4:47.0% 7/8:46.8% 8/4:46.6% 8/8:46.5%
%%svatava:er: showwer_nist test_symmetric_rerun/*/*/*0.0.ctm.filt.dtl
%%3/4:47.3% 3/8:47.0% 4/4:46.5% 4/8:46.4% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.8% 7/4:46.9% 7/8:45.0% 8/4:45.0% 8/8:45.1%

%%3/4:47.3% 3/8:47.1% 4/4:46.6% 4/8:46.3% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.9%

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|rcccccc|}           \hline 
  GMM:            &  \multicolumn{6}{c|}{    52.5                          }   \\ \hline  % from csl paper
                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
                  &  2700  &  4k    &  6k  & 9k   &  12k   &  16k    \\  \hline
% in /homes/eva/q/qghoshal/work/a3.0
% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  SGMM:             & 48.8   & 48.2 & 48.0  & 47.7  &{\bf 47.4}   &  47.5  \\
% in /homes/eva/q/qghoshal/work/a3.1
% showwer_nist t05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  +spk-vecs:        & 47.6   & 47.0  & 46.4  & 46.4  & 46.1  &{\bf  45.9} \\
% showwer_nist 05_test_sgmm+spkvecs+wts/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
  +symmetric,exact: & 46.3    & 45.6 &  45.2  & 44.8  & 44.5 &{\bf 44.4}  \\ 
% ./bin/showwer_nist
%05_test_sgmm+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
%3/8:46.5% 4/8:45.6% 5/8:45.0% 6/8:44.6% 7/8:44.4%
+symmetric,inexact & 46.5    & 45.6  &  45.0  & 44.6  & {\bf 44.4}&           \\ \hline
  \end{tabular}
   \caption{CallHome English: WERs without CMLLR adaptation }
% \vspace*{-0.2in}
 \label{tab:callhome:nocmllr}
\end{table}

Table~\ref{tab:callhome:nocmllr} shows experiments without CMLLR adaptation; the only
normalization is cepstral mean and variance normalization.  Using the symmetric
model reduced WER from 45.9\% to 44.4\%, a 1.5\% absolute improvement.   The inexact
update gave the same improvement as the exact update.

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|cccccc|}           \hline 
  GMM:            &  \multicolumn{6}{c|}{    49.7                             }   \\ 
  +SAT:           &  \multicolumn{6}{c|}{    46.0                             }   \\ \hline 
                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
                  &  2700  &  4k    &  6k   & 9k     &  12k       &  16k    \\  \cline{2-7}
% in /homes/eva/q/qghoshal/work/a3.0
% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  SGMM+spk-vecs:  & 46.5    & 45.5  & 45.2   & 45.4  & 44.8        & {\bf 44.7} \\
 +symmetric,exact & 44.9    & 44.4  &  44.1  & 43.2  & {\bf 42.8}  & 42.9      \\ 
%./bin/showwer_nist
%06_test_cmllr+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
+symmetric,inexact& 45.2    & 44.1  &  43.5  &  43.4  & {\bf 43.3} &           \\ \hline
  \end{tabular}
   \caption{CallHome English: WERs with CMLLR adaptation}
% \vspace*{-0.2in}
 \label{tab:callhome:cmllr}
\end{table}

Table~\ref{tab:callhome:cmllr} shows experiments with CMLLR adaptation.  The exact
update gives 1.9\% absolute improvement and the inexact update gives 1.4\% absolute
improvement.  Note that these are the same models as the previous table, tested with
CMLLR, and we attribute the difference between the exact and inexact models on this 
setup to statistical noise; further experiments will have to tell us whether, in general,
there is a difference between the exact and inexact updates.

%%Note that Speaker Adaptive Training (SAT) gives a large improvement for the
%%baseline GMM-based model, but we do not do this for our model: as reported
%%in~\cite{ubmpaper}, we did not get an improvement from SAT with the SGMM; we
%%are unsure why SAT gave an unusually large gain for the baseline model.

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|ccccccc|}           \hline 
 \multirow{2}{*}{GMM}
            &     \multicolumn{7}{c|}{\#Gauss per state}     \\
            &   20  &  26    &   32  &  34    &   36    &   38   & 40   \\  \cline{2-8}
 -          &       & 36.8   & 36.6  & 36.4   &  36.4   &  36.4  & 36.4  \\  
 CMLLR      &       & 34.8   & 34.5  & 34.4   & 34.3    & 34.5   & 34.3  \\
 STC        &       & 35.4   & 35.3  &        &  35.2   &        &      \\
    +CMLLR  &       & 33.1   & 32.9  &        &  32.9   &        &      \\ \hline
 \multirow{2}{*}{SGMM} 
            &    \multicolumn{7}{c|}{\#Substates}                      \\
            & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
 unadapted  &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ 
  CMLLR     &       &       &  32.2  &        &        &        &       \\
 +spk-vecs  &  32.0 & 31.7  &  31.4  & 31.2   & 30.8   &        &       \\ 
\hspace{-0.08in}
 +symmetric\hspace{-0.1in} &  31.9& 31.7   & 31.3   & 31.0   & 30.6   &        &       \\ \hline
  \end{tabular}
   \caption{Switchboard: WERs, with VTLN}
% \vspace*{-0.2in}
 \label{tab:switchboard:vtln}
\end{table}

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|cccccc|}           \hline 
 \multirow{2}{*}{GMM}
            &     \multicolumn{6}{c|}{\#Gauss per state}     \\
            &       &        &       &        &   36    &         \\  \cline{2-7}
 -          &       &        &       &        &  39.2   &        \\  
 CMLLR      &       &        &       &        &  37.0   &        \\
 STC        &       &        &       &        &  38.0   &        \\
 +CMLLR     &       &        &       &        &  35.2   &         \\ \hline
 \multirow{2}{*}{SGMM} 
            &    \multicolumn{6}{c|}{\#Substates}                      \\
            & 30k   &  40k  &   50k  &  75k   &  100k  &         \\  \cline{2-7}
 unadapted  &  37.9 & 37.5  &  37.1  & 36.6   &  36.3  &          \\ \hline
CMLLR+spk-vecs&33.9 & 33.5  &  33.4  &        &        &          \\
 +symmetric & 33.8  & 33.0  &  33.2  &        &        &          \\ \hline
  \end{tabular}
   \caption{Switchboard: WERs, no VTLN}
% \vspace*{-0.2in}
 \label{tab:switchboard:novtln}
\end{table}

%% \begin{table}[h]
%%  \renewcommand\arraycolsep{-0.1in}
%%   \begin{tabular}{|l|ccccccc|}           \hline 
%%             &     \multicolumn{7}{c|}{\#Gauss per state}     \\
%%             &  16   &   26   &  32   &   36   & 40       &      &    \\  \cline{2-8}
%%   GMM:      &  37.8 & 36.8   & 36.6  & 36.4   &  36.4    &      &     \\  
%% +STC:       &       &        & 35.6  &        &          &       &     \\ \hline
%%             &    \multicolumn{7}{c|}{\#Substates}                      \\
%%             & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
%%   SGMM:     &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ \hline
%%   \end{tabular}
%%    \caption{Switchboard: WERs, unadapted (except for VTLN)}
%% % \vspace*{-0.2in}
%%  \label{tab:switchboard:unadapted}
%% \end{table}

Next we discuss Switchboard experiments.  Our Switchboard system was trained on
278 hours of data from Switchboard I and II, and CallHome English.  Models were
tested on the Hub5 Eval01 test set (just over 6 hours long).  We used PLP
features with cepstral mean and variance normalization, and Vocal Tract Length
Normalization (VTLN).  The bigram language model used during decoding was taken
from the AMI RT'07 system described in~\cite{NIST:2007}; we used a recognition
lexicon of 50K words.  Our baseline GMM models were built with
HTK~\cite{htkbook}.  Tables~\ref{tab:switchboard:vtln}
and~\ref{tab:switchboard:novtln} show results with VTLN, and without VTLN,
respectively.  We did the baseline experiments with Constrained MLLR (CMLLR; a.k.a. fMLLR), 
and Semi-tied Covariance (STC; a.k.a. MLLT).  With the SGMMs, 
we used the exact update for the $\u_i$ quantities in the symmetric case.  
In both cases the symmetric extension to the model gives a much smaller
improvement than on the CallHome setup.  We are not sure of the reason for this.
Note that we do not show Speaker Adapted Training (SAT) results for the GMM baseline
because we did not see an improvement (we tried SAT after STC which would anyway
reduce the gains).

%%Table~\ref{tab:switchboard:unadapted}
%%shows experiments without adaptation except for VTLN (i.e. no ``speaker vectors'' or
%%CMLLR).  The GMM line with ``+STC'' includes a global semi-tied covariance transform (STC),
%%also known as a Maximum Likelihood Linear Transform (MLLT).
%%The SGMM is 1.9\% absolute (5.3\% relative) better than the best GMM baseline.

%% \begin{table}[h]
%%  \renewcommand\arraycolsep{-0.1in}
%%   \begin{tabular}{|l|ccccc|}           \hline 
%%                 &     \multicolumn{5}{c|}{\#Gauss per state}     \\
%%                 &  16   &   26   &  32   &   36   &   40       \\  \cline{2-6}
%%       GMM:      &       & 34.8   & 34.5  & 34.3   & 34.3       \\  \hline
%%      +STC:      &       &        & 34.1  &        &           \\  \hline
%%                 &    \multicolumn{5}{c|}{\#Substates}                      \\
%%                 & 30k   &  40k  &   50k  &  75k   &  100k   \\  \cline{2-6}
%%  SGMM+spk-vecs: &  33.0 & 32.5  &  32.0  & 31.7   & 31.4     \\ 
%%  +symmetric:    &  32.9 & 32.3  &  31.9  & 31.7   & 31.3     \\ \hline
%%   \end{tabular}
%%    \caption{Switchboard: WERs, adapted (all with VTLN+CMLLR)}
%% % \vspace*{-0.2in}
%%  \label{tab:switchboard:adapted}
%% \end{table}

%%Table~\ref{tab:switchboard:adapted} shows adapted results on Switchboard; these all include
%%CMLLR, and the SGMM systems also use speaker vectors.  Comparing the best GMM number with the best SGMM
%%number (with the symmetric model), we get 2.8\% absolute improvement (8.2\% relative).  Based on
%%other experiments on a slightly different setup, we believe that Speaker Adapative Training would
%%improve the baseline by about 0.3\% absolute.

%% The unexpected thing in Table~\ref{tab:switchboard:adapted} is that the symmetric modification gave 
%% us almost no improvement.  Although it is common to get slightly different results on different
%% datasets, we are not accustomed to seeing such large differences.

Table~\ref{tab:likes} compares
the average acoustic likelihood in the three passes of decoding, which reveals the effect of the
symmetric modification on the likelihood.  The ``baseline'' rows are the SGMM with speaker vectors 
and CMLLR, but without the symmetric modification.  As expected, the likelihood before adaptation
is slightly worse (because there is a mismatch between the models, which were adaptively trained,
and the data), but it gets better after adaptation.  In both cases the likelihood improvement per
frame, after adaptation, was about 0.1 (in natural-logarithm units).  This makes it 
hard to interpret the differences in results between the CallHome and Switchboard setups, because
the effect on the likelihoods is so similar.  We intend to do further experiments on other
data-sets to find which results are more typical.

% see ~/er/RESULTS.icassp11 on BUT systems.
\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|cr|ccc|}           \hline 
           &              &   \multicolumn{3}{c|}{ Decoding pass }           \\
           &              &   1         &     2       &          3            \\  
           &              & {\footnotesize (no-adapt)}  &{\footnotesize  +spk-vecs}  & {\footnotesize +CMLLR}         \\  \hline
  Call     &{\footnotesize SGMM+spk-vecs}&  -65.44     &   -63.62    &  -62.56              \\ 
  Home     & {\footnotesize +symmetric}    &  -65.57     &   -63.50    &  -62.45            \\ \hline
  Switch-  & {\footnotesize SGMM+spk-vecs}&  -60.07     &  -57.78     &  -58.86             \\
  board    & {\footnotesize +symmetric}    &  -60.17     &  -57.68     &  -56.76            \\ \hline
  \end{tabular}
   \caption{Acoustic likelihoods on the three test-time decoding passes}
% \vspace*{-0.2in}
 \label{tab:likes}
\end{table}

%% Baseline:
%% merlin:er: grep Per test_arnab*rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/\
%% $tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -65.441097461703 over 655612 frames
%% pass 2: -63.6198843852782 over 655612 frames
%% pass 3: -62.5628536429138 over 655612 frames

%% Experiment:
%% merlin:er: grep Per test_symmetric_rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{\
%% $n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -65.5657124102808 over 655612 frames
%% pass 2: -63.5015519090499 over 655612 frames
%% pass 3: -62.4537370786898 over 655612 frames
%%  Martin's setup:
%% cd /mnt/matylda3/karafiat/SGMM/exp/ctstrain04.v0.xwrd.sn.4.vtln

%% Baseline:
%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=split\
%% (`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -60.0675595633454 over 2228237 frames
%% pass 2: -57.7841323235589 over 2228237 frames
%% pass 3: -56.8589936376831 over 2228237 frames

%% Experiment:
%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_sym_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=s\
%% plit(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -60.1726961158082 over 2228237 frames
%% pass 2: -57.681438551532 over 2228237 frames
%% pass 3: -56.7608517686853 over 2228237 frames

@@

We have described a modification to the Subspace Gaussian Mixture Model which we
call the Symmetric SGMM.  This is a very natural extension which removes an
asymmetry in the way the Gaussian mixture weights were previously computed.  The
extra computation is minimal but the memory used for the acoustic model is nearly
doubled.  Our experimental results were inconsistent: on one setup we got a large
improvement of 1.5\% absolute, and on another setup it was much smaller.

We would also like to report our progress on releasing open-source software that
supports the SGMM modeling approach.  An official announcement, with additional
co-authors, will follow within the next year.  We are developing an open-source
(Apache-licensed) C++ speech recognition toolkit that uses the OpenFst
library~\cite{openfst}.  Most aspects of the toolkit are not related directly to
SGMMs, but SGMMs will be one of the acoustic models the toolkit natively
supports.  Most likely the toolkit will already have been released by the time this 
is published.

>>>>

@@

Conclusions here
